\textsc{HubbleCLIP}: Associating Astronomical Observations and Natural Language with Multi-Modal Models

### 2024-01-03 paper/main.tex
Bib

\subsection{Data Selection and Pre-processing}

@@

CLIP \citep[Contrastive Language-Image Pretraining;][]{radford2021learning} is a multi-modal model pre-trained on a large corpus of image-text pairs via weak supervision using a contrastive loss. Given a minibatch $\mathcal{B}$ of $|\mathcal{B}|$ image-text pairs $\{(I_i, T_i)\}$, the goal is to align the learned representations of corresponding (positive) pairs $(I_i, T_i)$ while repelling the representations of unaligned (negative) pairs $(I_i, T_{j\neq i})$. Image and text encoders $f(\cdot)$ and $g(\cdot)$ are used to map images and text to a common embedding space. The standard softmax-based bidirectional variation of the InfoNCE~\citep{oord2018representation} contrastive loss function, as used by CLIP, is particularly effective for multimodal learning. This bidirectionality is crucial as it ensures the model learns to map both images to text and text to images with equal importance. This symmetry in learning is essential for tasks that require a mutual understanding and interchangeability between visual and textual representations, such as image captioning, text-to-image synthesis, and cross-modal retrieval. The bidirectional loss is given by \citep{radford2021learning}
\begin{equation}
\mathcal L(\mathcal B) = -\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\overbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_i \cdot {y}_j / \tau}}}^{\text {image } \rightarrow \text { text softmax }}+\overbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_j \cdot {y}_i / \tau}}}^{\text {text } \rightarrow \text { image softmax }}\right)
\end{equation}
where ${x}_i=\frac{f\left(I_i\right)}{\left\|f\left(I_i\right)\right\|_2}$ and ${y}_i=\frac{g\left(T_i\right)}{\left\|g\left(T_i\right)\right\|_2}$ are the normalized representations of the $i$-th image and text, respectively, and $\tau$ is a (typically learnable) temperature hyperparameter. This bidirectional approach not only enhances the model's ability to create robust and versatile embeddings but also aligns with the intuitive understanding of how humans interact with and interpret multimodal information, further bridging the gap between artificial and natural intelligence in multimodal contexts.

As the base model, we use the CLIP-ViT-B/16 \citep{radford2021learning} variant trained by OpenAI. This model uses a (12-layer, 12-head, 768-embedding dimension) vision transformer as the image encoder and a (12-layer, 8-head, 512-embedding dimension) sequence transformer as the text backbone. The text encoder has a maximum length of 77-tokens and the image encoder a native resolution of $224\times224$. Linear projection layers map the outputs of the image and text encoders to a common embedding space of dimension 512. In total, the model has 149,620,737 trainable parameters. The model was trained on 400 million image-text pairs from the internet.

@@

The model is trained over 50,000 steps with 5000 linear warmup steps and cosine decay using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with peak learning rate of either $10^{-5}$ or $10^{-6}$ and weight decay $10^{-3}$. Training takes approximately 6 hours on 4 Nvidia A100 GPUs.

@@

@@

\subsubsection*{Acknowledgments}
This work is supported by the National Science Foundation under Cooperative Agreement PHY-2019786 (The NSF AI Institute for Artificial Intelligence and Fundamental Interactions, \url{http://iaifi.org/}). This material is based upon work supported by the U.S. Department of Energy, Office of Science, Office of High Energy Physics of U.S. Department of Energy under grant Contract Number  DE-SC0012567. The computations in this paper were run on the FASRC Cannon cluster supported by the FAS Division of Science Research Computing Group at Harvard University.

@@

\appendix
\section{Summarization via regex-guided generation}
\label{app:summarization}

@@

>>>>

\subsection{Data Selection and Pre-Processing}

@@

CLIP \citep[Contrastive Language-Image Pretraining;][]{radford2021learning} is a multi-modal model pre-trained on a large corpus of image-text pairs via weak supervision using a contrastive loss. Given a minibatch $\mathcal{B}$ of $|\mathcal{B}|$ image-text pairs $\{(I_i, T_i)\}$, the goal is to align the learned representations of corresponding (positive) pairs $(I_i, T_i)$ while repelling the representations of unaligned (negative) pairs $(I_i, T_{j\neq i})$. Image and text encoders $f(\cdot)$ and $g(\cdot)$ are used to map images and text to a common embedding space. The standard softmax-based bidirectional variation of the InfoNCE~\citep{oord2018representation} contrastive loss function, as used by CLIP, is particularly effective for multimodal learning. This bidirectionality is crucial as it ensures the model learns to map both images to text and text to images with equal importance. This symmetry in learning is essential for tasks that require a mutual understanding and interchangeability between visual and textual representations, such as image captioning, text-to-image synthesis, and cross-modal retrieval. The bidirectional loss is given by \citep{radford2021learning}
\begin{equation}
\mathcal L(\mathcal B) = -\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\overbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_i \cdot {y}_j / \tau}}}^{\text {image } \rightarrow \text { text softmax }}+\overbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_j \cdot {y}_i / \tau}}}^{\text {text } \rightarrow \text { image softmax }}\right)
\end{equation}
where ${x}_i=\frac{f\left(I_i\right)}{\left\|f\left(I_i\right)\right\|_2}$ and ${y}_i=\frac{g\left(T_i\right)}{\left\|g\left(T_i\right)\right\|_2}$ are the normalized representations of the $i$-th image and text, respectively, and $\tau$ is a (typically learnable) temperature hyperparameter. 
% This bidirectional approach not only enhances the model's ability to create robust and versatile embeddings but also aligns with the intuitive understanding of how humans interact with and interpret multimodal information, further bridging the gap between artificial and natural intelligence in multimodal contexts.

As the base model, we use the CLIP-ViT-B/16 \citep{radford2021learning} variant trained by OpenAI. This model uses a (12-layer, 12-head, 768-embedding dimension) vision transformer as the image encoder and a (12-layer, 8-head, 512-embedding dimension) sequence transformer as the text backbone. The text encoder has a maximum length of 77 tokens and the image encoder a native resolution of $224\times224$. Linear projection layers map the outputs of the image and text encoders to a common embedding space of dimension 512. In total, the model has 149,620,737 trainable parameters. The model was trained on 400 million image-text pairs from the internet.

@@

Models were instantiated and trained using the \package{Jax} \citep{jax2018github} ecosystem. The model is trained over 50,000 steps with 5000 linear warmup steps and cosine decay using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with peak learning rate of either $10^{-5}$ or $10^{-6}$ and weight decay $10^{-3}$. Training takes approximately 6 hours on 4 Nvidia A100 GPUs.

@@

The code, dataset, and models used in this work is available at \url{https://www.github.com/smsharma/HubbleCLIP}.

This work relied on the \package{Astroquery} \citep{2019AJ....157...98G}, \package{bitsandbytes} \citep{dettmers2022llmint8}, \package{Flax} \citep{flax2020github}, \package{Jax} \citep{jax2018github}, \package{Jupyter} \citep{Kluyver2016jupyter}, \package{Matplotlib} \citep{Hunter:2007}, \package{Numpy} \citep{harris2020array}, \package{Optax} \citep{deepmind2020jax}, \package{Outlines}, \package{Pandas}, \package{Pydantic}, \package{PyTorch} \citep{paszke2019pytorch}, \package{SciPy} \citep{2020SciPy-NMeth}, \package{Transformers} \citep{wolf2019huggingface}, and \package{wandb} \citep{wandb} software packages.

@@

\subsubsection*{Acknowledgments}
This work is supported by the National Science Foundation under Cooperative Agreement PHY-2019786 (The NSF AI Institute for Artificial Intelligence and Fundamental Interactions, \url{http://iaifi.org/}). This material is based upon work supported by the U.S. Department of Energy, Office of Science, Office of High Energy Physics of U.S. Department of Energy under grant Contract Number  DE-SC0012567. YS was supported by the Research Science Institute (RSI) program at MIT. The computations in this paper were run on the FASRC Cannon cluster supported by the FAS Division of Science Research Computing Group at Harvard University.

@@

\appendix
\section{Summarization via Regex-Guided Generation}
\label{app:summarization}

@@

\section{Additional Evaluation Metrics and Ablations}
\label{app:ablations}
